\bf Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2024-07-31 paper.qmd
Changes to Figure 8

The values $(s_1, \ s_2)$ define the position of the centroid of the bottom left hexagon. By default, this is at $s_1 = -q, s_2 = -qr_2$, where $q$ is the buffer sound the data. The choice of these values can have some effect on the distribution of bin counts. FIG XXX(a) illustrates this. The distribution of bin counts for $s_1$ varying between $-0.1-0.0$ is shown. Generally, a more uniform distribution among these possibilities would indicate that the bins are reliably capturing the underlying distribution of observations.

@@

mse_scurve_b <- ggplot(error_scurve, 
                     aes(x = b, 
                         y = log(MSE),
                         color = buff,
                         group = buff)) + 
  geom_point(size = 1) +
  geom_line(linewidth = 0.05) + 
  geom_vline(xintercept = 220, linetype="solid", 
             color = "black", linewidth=0.1) +
  geom_vline(xintercept = 112, linetype=2, 
             color = "black", linewidth=0.1) +
  geom_vline(xintercept = 364, linetype=2, 
             color = "black", linewidth=0.1) +
  scale_colour_continuous_sequential(palette = "Magenta") +
  labs(x = expression(b), y = "log(MSE)", color = expression(q)) +
  ggtitle("(a)") +
  theme(aspect.ratio = 0.75,
        axis.text.x = element_text(size = 12),
        axis.text.y = element_text(size = 12),
        axis.title.x = element_text(size = 12),
        axis.title.y = element_text(size = 12, 
            angle=90),
        plot.title = element_text(size = 12),
        legend.title = element_text(size = 12),
        legend.text = element_text(size = 10),
        legend.key.height = unit(1, 'cm'),
        legend.key.width = unit(1, 'cm'),
        legend.key.size = unit(1, 'cm'))
```

@@

By default, when assessing the choice of $b_1$, the total number of bins is measured by the number of **non-empty** bins. This more accurately reflects the hexagon grid relative the MSE than the full number of bins in the grid. It may also be beneficial to remove low count bins also, in the situation where data is clustered or stringy, where the observed data is sparse. In order to decide if this is necessary, you would examine the distribution of bin counts, or the density which puts the counts on a standard scale. If there is something of a gap at low values, this would suggest a potential value to use as a cutoff. Alternatively, one could choose to remove based on a percentile, the bins with density in the lowest 5% of all bins, for example. FIG XXXc illustrates the effect on the model representation of removing bins below different percentages. Generally, we would urge caution in removing low count bins.

@@

cell_count_scurve <- ggplot(df_bin_centroids_all, 
                            aes(x = reorder(as.factor(hexID), 
                                            -bin_counts), 
                                y = bin_counts,
                                group = a1,
                                color = a1)) + 
  geom_line() + 
  # geom_hline(yintercept = benchmark1, color = "black", linewidth=0.5, alpha = 0.5) +
  scale_colour_continuous_sequential(palette = "Magenta") +
  xlab("hexagon id (re-ordered)") + 
  ylab("standardized bin count") +
  ggtitle("(b)") +
  theme(aspect.ratio = 0.75,
        axis.title.x = element_text(size = 12),
        axis.title.y = element_text(size = 12, angle = 90),
        axis.text.x = element_blank(),
        axis.text.y = element_text(size = 12),
        plot.title = element_text(size = 12))
```

@@

The decision on edge length removal is made based on the distribution of edge lengths. In particular, a gap between values, where there a concentration of small values and then a few larger values, likely suggests a cutoff for edge removal. Because the triangulation is typically done on the hexagon centroids, there are particular discrete edge lengths, based on bin widths. @fig-NLDR-scurve XXX illustrates edge length distributions.

@@

mse_scurve_b + cell_count_scurve + 
mse_scurve_lwd + 
distance_hist + plot_layout(ncol=2)
```

@@

XXXX I think we can remove the error figure, it's not part of the story

>>>>

The values $(s_1, \ s_2)$ define the position of the centroid of the bottom left hexagon. By default, this is at $s_1 = -q, s_2 = -qr_2$, where $q$ is the buffer sound the data. The choice of these values can have some effect on the distribution of bin counts. @fig-param-scurve (a) illustrates this. The distribution of bin counts for $s_1$ varying between $-0.1-0.0$ is shown. Generally, a more uniform distribution among these possibilities would indicate that the bins are reliably capturing the underlying distribution of observations.

@@

# mse_scurve_b <- ggplot(error_scurve, 
#                      aes(x = b, 
#                          y = log(MSE),
#                          color = buff,
#                          group = buff)) + 
#   geom_point(size = 1) +
#   geom_line(linewidth = 0.05) + 
#   geom_vline(xintercept = 220, linetype="solid", 
#              color = "black", linewidth=0.1) +
#   geom_vline(xintercept = 112, linetype=2, 
#              color = "black", linewidth=0.1) +
#   geom_vline(xintercept = 364, linetype=2, 
#              color = "black", linewidth=0.1) +
#   scale_colour_continuous_sequential(palette = "Magenta") +
#   labs(x = expression(b), y = "log(MSE)", color = expression(q)) +
#   ggtitle("(a)") +
#   theme(aspect.ratio = 0.75,
#         axis.text.x = element_text(size = 12),
#         axis.text.y = element_text(size = 12),
#         axis.title.x = element_text(size = 12),
#         axis.title.y = element_text(size = 12, 
#             angle=90),
#         plot.title = element_text(size = 12),
#         legend.title = element_text(size = 12),
#         legend.text = element_text(size = 10),
#         legend.key.height = unit(1, 'cm'),
#         legend.key.width = unit(1, 'cm'),
#         legend.key.size = unit(1, 'cm'))

mse_scurve_b <- ggplot(error_scurve, 
                     aes(x = b, 
                         y = MSE,
                         color = as.factor(buff),
                         group = as.factor(buff))) + 
  geom_point(size = 1) +
  geom_line(linewidth = 0.05) + 
  # geom_vline(xintercept = 220, linetype="solid", 
  #            color = "black", linewidth=0.1) +
  # geom_vline(xintercept = 112, linetype=2, 
  #            color = "black", linewidth=0.1) +
  # geom_vline(xintercept = 364, linetype=2, 
  #            color = "black", linewidth=0.1) +
  scale_colour_discrete_qualitative(palette = "Dark 2") +
  labs(x = expression(b), y = "MSE", color = expression(q)) +
  ggtitle("(b)") +
  theme(aspect.ratio = 0.75,
        axis.text.x = element_text(size = 12),
        axis.text.y = element_text(size = 12),
        axis.title.x = element_text(size = 12),
        axis.title.y = element_text(size = 12, 
            angle=90),
        plot.title = element_text(size = 12),
        legend.title = element_text(size = 12),
        legend.text = element_text(size = 10),
        legend.key.height = unit(1, 'cm'),
        legend.key.width = unit(1, 'cm'),
        legend.key.size = unit(1, 'cm'))
```

@@

By default, when assessing the choice of $b_1$, the total number of bins is measured by the number of **non-empty** bins. This more accurately reflects the hexagon grid relative the MSE than the full number of bins in the grid. It may also be beneficial to remove low count bins also, in the situation where data is clustered or stringy, where the observed data is sparse. In order to decide if this is necessary, you would examine the distribution of bin counts, or the density which puts the counts on a standard scale. If there is something of a gap at low values, this would suggest a potential value to use as a cutoff. Alternatively, one could choose to remove based on a percentile, the bins with density in the lowest 5% of all bins, for example. @fig-param-scurve (c) illustrates the effect on the model representation of removing bins below different percentages. Generally, we would urge caution in removing low count bins.

@@

cell_count_scurve <- ggplot(df_bin_centroids_all, 
                            aes(x = reorder(as.factor(hexID), 
                                            -bin_counts), 
                                y = bin_counts,
                                group = as.factor(a1),
                                color = as.factor(a1))) + 
  geom_line() + 
  # geom_hline(yintercept = benchmark1, color = "black", linewidth=0.5, alpha = 0.5) +
  scale_colour_discrete_qualitative(palette = "Dark 2") +
  xlab("hexagon id (re-ordered)") + 
  ylab("bin count") +
  ggtitle("(a)") +
  theme(aspect.ratio = 0.75,
        axis.title.x = element_text(size = 12),
        axis.title.y = element_text(size = 12, angle = 90),
        axis.text.x = element_blank(),
        axis.text.y = element_text(size = 12),
        plot.title = element_text(size = 12))
```

@@

The decision on edge length removal is made based on the distribution of edge lengths. In particular, a gap between values, where there a concentration of small values and then a few larger values, likely suggests a cutoff for edge removal. Because the triangulation is typically done on the hexagon centroids, there are particular discrete edge lengths, based on bin widths. @fig-NLDR-scurve (d) illustrates edge length distributions.

@@

cell_count_scurve + mse_scurve_b +  
mse_scurve_lwd + 
distance_hist + plot_layout(ncol=2)
```

@@

### 2024-07-31 paper.tex
Changes to Figure 8

The values \((s_1, \ s_2)\) define the position of the centroid of the
bottom left hexagon. By default, this is at \(s_1 = -q, s_2 = -qr_2\),
where \(q\) is the buffer sound the data. The choice of these values can
have some effect on the distribution of bin counts. FIG XXX(a)
illustrates this. The distribution of bin counts for \(s_1\) varying
between \(-0.1-0.0\) is shown. Generally, a more uniform distribution
among these possibilities would indicate that the bins are reliably
capturing the underlying distribution of observations.

@@

By default, when assessing the choice of \(b_1\), the total number of
bins is measured by the number of \textbf{non-empty} bins. This more
accurately reflects the hexagon grid relative the MSE than the full
number of bins in the grid. It may also be beneficial to remove low
count bins also, in the situation where data is clustered or stringy,
where the observed data is sparse. In order to decide if this is
necessary, you would examine the distribution of bin counts, or the
density which puts the counts on a standard scale. If there is something
of a gap at low values, this would suggest a potential value to use as a
cutoff. Alternatively, one could choose to remove based on a percentile,
the bins with density in the lowest 5\% of all bins, for example. FIG
XXXc illustrates the effect on the model representation of removing bins
below different percentages. Generally, we would urge caution in
removing low count bins.

@@

The decision on edge length removal is made based on the distribution of
edge lengths. In particular, a gap between values, where there a
concentration of small values and then a few larger values, likely
suggests a cutoff for edge removal. Because the triangulation is
typically done on the hexagon centroids, there are particular discrete
edge lengths, based on bin widths. Figure~\ref{fig-NLDR-scurve} XXX
illustrates edge length distributions.

>>>>

The values \((s_1, \ s_2)\) define the position of the centroid of the
bottom left hexagon. By default, this is at \(s_1 = -q, s_2 = -qr_2\),
where \(q\) is the buffer sound the data. The choice of these values can
have some effect on the distribution of bin counts.
Figure~\ref{fig-param-scurve} (a) illustrates this. The distribution of
bin counts for \(s_1\) varying between \(-0.1-0.0\) is shown. Generally,
a more uniform distribution among these possibilities would indicate
that the bins are reliably capturing the underlying distribution of
observations.

@@

By default, when assessing the choice of \(b_1\), the total number of
bins is measured by the number of \textbf{non-empty} bins. This more
accurately reflects the hexagon grid relative the MSE than the full
number of bins in the grid. It may also be beneficial to remove low
count bins also, in the situation where data is clustered or stringy,
where the observed data is sparse. In order to decide if this is
necessary, you would examine the distribution of bin counts, or the
density which puts the counts on a standard scale. If there is something
of a gap at low values, this would suggest a potential value to use as a
cutoff. Alternatively, one could choose to remove based on a percentile,
the bins with density in the lowest 5\% of all bins, for example.
Figure~\ref{fig-param-scurve} (c) illustrates the effect on the model
representation of removing bins below different percentages. Generally,
we would urge caution in removing low count bins.

@@

The decision on edge length removal is made based on the distribution of
edge lengths. In particular, a gap between values, where there a
concentration of small values and then a few larger values, likely
suggests a cutoff for edge removal. Because the triangulation is
typically done on the hexagon centroids, there are particular discrete
edge lengths, based on bin widths. Figure~\ref{fig-NLDR-scurve} (d)
illustrates edge length distributions.
